Unshuffling Permutations

### 2015-09-11 article.tex
reduction

  \begin{proof}[of Claim~\ref{claim:one (nu_1, nu'_1)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Then it follows that there exists an edge $(x, y)$ in $\mathcal{M}$,
    $1 \leq x \leq N_1$, that is neither a
    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
    (In other words, $x \leq N_1$ and $y > 2N_1 + N_2$.)
    Therefore, since $\mathcal{M}$ is containment-free,
    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
    in $\mathcal{M}$.
    Then it follows that every position $x$ of $\nu'_1$
    (\emph{i.e.}, $N_1+N_2 < x \leq 2N_1+N_2$)
    is involved in an edge $(x, y)$ of $\mathcal{M}$ with $y \geq 2N_1+N_2$.
    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
    Then it follows that $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
    We first prove that there exists at most one
    $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
    Indeed, suppose, aiming at a contradiction, that there exist
    two $(\nu_2, \nu'_1)$-edges $e = (a, b)$ and $e' = (a', b')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free,
    we may assume that $a < a'$ and $b < b'$.
    Then it follows that $\mu(a) > \mu(a')$
    (since $\nu_2$ is decreasing) and $\mu(b) < \mu(b')$
    (since $\nu'_1$ is increasing),
    which contradicts our hypothesis.

    Suppose now, still aiming at a contradiction, that there exists
    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, b)$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge}, there exists at least
    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Furtheremore, since
    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
    every position in $\nu_1$ is involved either in a
    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
    (otherwise $\mathcal{M}$ would not be containment-free).
    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
    we conclude that there exists a position, say $x$, in $\nu'_1$
    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
    nor a $(\nu'_1, \nu'_1)$-edge.
    Write $(x, y)$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $y > 2N_1 + N_2$, this edge.
    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
    and $\mu(b) > \mu(y)$ (since $\nu'_1$ is above any other gadget),
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:one (nu_2, nu'_2)-edge}]
    Suppose, aiming at a contradiction there there is no
    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
    Notice that there is neither
    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
    in $\mathcal{M}$.
    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k$.
    Hence $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}]
    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
    $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}).
    \qed
  \end{proof}

  \begin{figure}[t!]
    \centering
      \begin{tikzpicture}
        [
          scale=0.25,
          label/.style={anchor=base}
        ]
        % nu_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu_3
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu'_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);

        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (x) at (2,4.5) {$x$};
        \node [label] (a) at (8,14.5) {$a$};
        \node [label] (y) at (14,-1.5) {$y$};
        \node [label] (b) at (16,-1.5) {$b$};
        \node (nu1) at (2.5,2.5) {$\nu_2$};
        \node (nu2) at (8.5,12.5) {$\nu_3$};
        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
        % crossings edges
        \draw [line width=1pt]
        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
        \draw [line width=1pt]
        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
      \end{tikzpicture}
    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
    }
  \end{figure}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
    Suppose, aiming at a contradiction, that there exists
    $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
    exists at least one $(\nu_2, \nu'_2)$-edge $(x, y)$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
    we have $y < b$.
    Therefore,
    we have $\mu(x) < \mu(a)$ (by construction) and
    $\mu(y) > \mu(b)$ (since $\nu'_2$ is decreasing),
    which contradicts our hypothesis.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}
    There is neither a $(\nu_1, \nu'_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}]
    Suppose, aiming at a contradiction, that there exists a
    $(\nu_1, \nu'_3)$-edge $(a, b)$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge} there exists
    $(\nu_1, \nu'_1)$-edge $(x, y)$ in $\mathcal{M}$,
    and ince $\mathcal{M}$ is containment-free we have $x < a$.
    Therefore, we have
    $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing)
    and
    $\mu(y) > \mu(b)$ (by construction),
    which contradicts our hypothesis.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.

    A similar argument shows that there is no
    $(\nu'_1, \nu_3)$-edge in $\mathcal{M}$.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
    Suppose, aiming at a contradiction, that there exist two
    $(\nu_2, \nu_3)$-edges $(a, b)$ and $(a', b')$ in $\mathcal{M}$.
    Since these two edges have to be crossing
    (Proposition~\ref{proposition:matching}), we may assume that
    $a < a'$ and $b < b'$.
    Therefore, we have
    $\mu(a) > \mu(a')$ (since $\nu_2$ is decreasing)
    and
    $\mu(b) < \mu(b')$ (since $\nu_3$ is increasing).
    According to Proposition~\ref{proposition:matching}, this is
    a contradiction.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:one (nu_3, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
    \qed
  \end{proof}

>>>>

  \begin{proof}[of Claim~\ref{claim:one (nu_1, nu'_1)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
    $1 \leq x \leq N_1$, that is neither a
    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
    Therefore, since $\mathcal{M}$ is containment-free,
    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
    in $\mathcal{M}$.
    Then it follows that every position $y$ of $\nu'_1$
    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
    (otherwise $\mathcal{M}$ would not be containment-free).
    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
    Then it follows that $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge}, there exists at least
    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Furtheremore, since
    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
    every position in $\nu_1$ is involved either in a
    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
    (otherwise $\mathcal{M}$ would not be containment-free).
    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
    we conclude that there exists a position, say $x$, in $\nu'_1$
    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
    nor a $(\nu'_1, \nu'_1)$-edge.
    Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
    and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:one (nu_2, nu'_2)-edge}]
    Suppose, aiming at a contradiction there there is no
    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
    Notice that there is neither
    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
    in $\mathcal{M}$.
    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k +4$.
    Hence $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}]
    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
    $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  \begin{figure}[t!]
    \centering
      \begin{tikzpicture}
        [
          scale=0.35,
          label/.style={anchor=base}
        ]
        % nu_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu_3
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu'_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);

        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (x) at (2,4.5) {$x$};
        \node [label] (a) at (8,14.5) {$a$};
        \node [label] (xp) at (14,-1.5) {$x'$};
        \node [label] (ap) at (16,-1.5) {$a'$};
        \node (nu1) at (2.5,2.5) {$\nu_2$};
        \node (nu2) at (8.5,12.5) {$\nu_3$};
        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
        % crossings edges
        \draw [line width=1pt]
        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
        \draw [line width=1pt]
        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
      \end{tikzpicture}
    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
    }
  \end{figure}

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
    Suppose, aiming at a contradiction, that there exists
    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
    we have $x' < a'$.
    Therefore,
    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
    which contradicts our hypothesis.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}
    There is neither a $(\nu_1, \nu_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}]
    Suppose, aiming at a contradiction, that there exists a
    $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge} there exists
    $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$,
    and since $\mathcal{M}$ is containment-free we have $x < a$.
    Therefore, we have
    $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing)
    and
    $\mu(x') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
    which contradicts our hypothesis.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.

    The same argument shows that there is no
    $(\nu'_1, \nu_3)$-edge in $\mathcal{M}$.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
    Apply Lemma~\ref{lemma:at most one edge monotone}.
    \qed
  \end{proof}

@@

  \begin{proof}[of Claim~\ref{claim:one (nu_3, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
    Combining Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge},
    Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge},
    Claim~\ref{claim:at most one (nu_2, nu_3)-edge},
    and Claim~\ref{claim:no (nu_3, nu'_2)-edge},
    we conclude that $N_3-1$ positions in $\nu_3$ are involved
    in edges of $\mathcal{M}$ that are
    neither ($\nu_1, \nu_3)$-edges,
    nor $(\nu_2, \nu_3)$-edges
    nor $(\nu'_1, \nu_3)$-edges
    nor $(\nu_3, \nu_3)$-edges.
    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
    $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis.
    \qed
  \end{proof}
  \begin{claim}
    \label{claim:no (sigma', nu_4)-edge}
    There is no $(\sigma', \nu_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (sigma', nu_4)-edge}]
    Combine Claim~\ref{claim:one (nu_3, nu'_3)-edge} with the fact that
    $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there exists
    a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
    in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free, we must have
    $x' < a'$.
    Therefore,
    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
    which contradicts our hypothesis.
    \qed
  \end{proof}

@@
